### 2022-07-06 paper/Manuscript.Rmd
Add empirical strategy

---
title: "Frame Competition"
author: "Nicolai Berk"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    fig_caption: yes
    includes:
      in_header: preamble.tex
subtitle: Manuscript prepared for the 12th Annual Conference of the European Political Science Association in Prague, Czech Republic
fontsize: 12pt
geometry: margin=1.5in
abstract: "Issue ownership theory assumes that the salience of specific issues is always more or less beneficial to a given party. I introduce the concept of 'frame competition' and argue that a given issue can be more or less beneficial for parties dependent on the dominant framing. I test this argument using data from the German 2017 Bundestag campaign, assessing the similarity of media coverage on migration to the parties' migration framing with topic vectors. Using VAR models, I show whether higher frame similarity is associated with better performance in the polls. The findings broaden our understanding party competition and show parties' capability to compete when the issue agenda is given."
bibliography: C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib
---

```{r setup, include = F}

knitr::opts_chunk$set(echo = F, message = F, warning = F, cache = F, fig.height=6, fig.width=10, fig.align='center', eval = T)

library(tidyverse)
library(dplyr)
library(data.table)
library(patchwork)
library(here)

# load self-made functions
for (filename in list.files(here("code/functions"))){
  source(here(paste0('code/functions/', filename)))
}

```

# Introduction

One of the most prominent explanations for parties' electoral success in scholarly and public debate is the issue agenda at a given election: the more prominent an issue was among political parties, news media, and voters, the more it would influence the latter's electoral decision making. Certain issues are therefor more favorable to given parties and parties will try to focus on 'their' issues in public communication [@Budge1983; @Petrocik1996; @Walgrave2012]. In extension, it is assumed that the main form of electoral competition in contemporary Western European democracies is *issue competition*, in the sense that parties compete to bring 'their' issue on the agenda [@GreenPedersen2019].

The present paper sets out to challenge the truism of 'issue ownership' by looking into the debate *within* issues. Building on a large body of experimental work, I argue that - dependent on the relative emphasis of different considerations in public debate, also known as *emphasis framing* - any issue can be more or less favorable to a given party. As issues on the media agenda usually emerge externally [@Boydstun2013], issue ownership theory suggests that parties' electoral fate is statically bound to the media agenda of the day. My model of *frame competition* gives parties substantial agency by suggesting that competition takes not only place about *which* issues are talked about, but also *how*.

I test this argument using data from the German 2017 Bundestag campaign, during which the migration issue - which became salient after the 2015 refugee movements - dominated public debate in the media and among parties, as well as voters' perception important problems in national politics. Using topic representations of media content and parties' manifestos as an anchor point, I track the similarity of the migration discourse in national newspapers to parties' communication about migration. Combining these daily similarity estimates with polling data from *Poll of Polls* in a vector auto-regression (VAR) model [@Sims1980], I assess whether parties' electoral prospects increase when the media discourse is more similar to a given party's representation of the migration issue. The findings potentially challenge the current model of party competition, underlining parties' agency even when the agenda is externally determined, and speak to the real-world implications of issue framing.

# Issue Ownership Theory

The classic formulation of issue ownership theory argues that parties 'own' issues when they are peceived as more competent handling the issue [@Budge1983]. If the problem pressure regarding that issue increases and the issue rises to prominence, voters are more likely to base their voting decision on that issue and choose the party that is considered most able to deal with the issue. Over time, scholars have differentiated between this 'competence-based' and an 'associative' form of issue ownership. The latter emphasizes the process of voters' implicit and spontaneous association of an issue with a given party, thereby increasing the likelihood to vote for that party [@Lefevere2015b].

Whichever the mechanism, both conceptualizations imply that parties perform better electorally when 'their' issue is on the top of the agenda, because voters either are more likely to perceive them as the most competent party in the current issue environment or because they are simply more likely to make it to the pool of options in a voters' mind:

*Issue Ownership Hypothesis: The more salient an issue is, the better the electoral prospect of the party owning the issue.*

As mentioned, issue ownership theory assumes that party competition is largely an effort of parties to get 'their' topics on the agenda. As citizens mostly perceive politics in the news media, this means that parties need to bring 'their' issues to the media's attention. However, existing research suggests that the news media rarely follows political actors' attempts to set the agenda, but rather 'leapfrogs' from issue to issue, usually because of external events [@Boydstun2013]. For example, the outbreak of a foreign war, a large movement of refugees, or environmental catastrophes - all events that political parties usually have limited to no control over - might come to dominate the news agenda despite parties' best efforts to talk about redistribution, law and order, or taxation. In this view, parties have ultimately very limited control over their electoral prospects.

# Framing and Party Competition

I suggest that issues can have a differential impact on parties electoral prospects, dependent on *how* they are discussed. Given an issue like migration dominates media coverage, this does not necessarily suggest that the issue owner - typically radical right parties - profits from this electorally. If, for example migration is discussed as a humanitarian issue, e.g. migrants dying at the border, or refugees fleeing war, the issue should favor those parties with more welcoming migration positions more compared to when migration is discussed as a security problem, emphasizing crime and the potential danger of terrorism.

Consider the example in figure \@ref(fig:afdmigration) below. The top panel shows the weekly share of news articles devoted to migration in the major German daily newspapers, from January 2015 to June 2016. As we can see, attention to migration experiences a huge increase in September 2015, when the decision was made to let refugees stranded at Keletti station in Budapest, Hungary, pass to Germany. Issue Ownership theory suggests that this large increase in salience should directly translate into electoral support for the issue owner - in this case, the radical right *AfD*.

```{r afdmigration, fig.height = 10}

# plot afd polling
de_polls <- 
  rjson::fromJSON(file = "https://www.politico.eu/wp-json/politico/v1/poll-of-polls/DE-parliament")$polls

de_polls_df <- data.frame()

for (row in de_polls){

  row <- as.data.frame(row)

  row$sample_size <- as.numeric(row$sample_size)

  de_polls_df <- 
    bind_rows(
      de_polls_df,
      row)

}

afd_poll_plot <- 
  de_polls_df %>% 
  mutate(date = as.Date(date)) %>% 
  filter(date < as.Date("2016-07-01"),
         date >= as.Date("2015-01-01")) %>% 
  mutate(week = lubridate::floor_date(date, "week")
         # , post = date >= as.Date("2016-01-01")
         ) %>% 
  group_by(week
           # , post
           ) %>% 
  summarise(parties.AfD = mean(parties.AfD)) %>% 
  ggplot(aes(week, parties.AfD
             # , group = post
             )) +
  geom_point() +
  geom_smooth(method = "gam") +
  # geom_vline(xintercept = as.Date("2015-09-01"), col = "red", lty = 2) +
  theme_minimal() +
  ylab("AfD vote share") +
  xlab("Date")

# plot media attention to migration - add spon if possible
mig_salience_plot <- 
  fread(here("data/media/bert_crime_daily.csv")) %>% 
  select(date_clean, paper, mig_share) %>% 
  mutate(date_new = lubridate::floor_date(date_clean, "week")) %>% 
  group_by(date_new, paper) %>% 
  summarise_all(mean) %>% 
  filter(date_new < as.Date('2016-07-01'),
         date_new >= as.Date("2015-01-01")) %>%
  ggplot(aes(x = date_new, y = mig_share, col = paper, lty = paper, shape = paper)) +
  geom_line() + geom_point() +
  # geom_vline(xintercept = as.Date("2015-09-01"), col = "red", lty = 2) +
  ggtitle('Migration salience in different newspapers') +
  xlab('') + ylab('Share of migration articles') +
  theme_minimal() +
  labs(color = "Outlet", lty = "Outlet", shape = "Outlet")

mig_salience_plot / afd_poll_plot

```

The AfD's electoral prospect as reported in opinion polls is plotted in the panel below, for the same time frame. As we can see, the increase in voting intention for the radical right party trails the increase in salience by several months. Why might that be? Is it really the case that it took voters three to six months to understand that migration was a pressing issue on the political agenda, despite the omnipresence of the issue in the news media? I suggest another explanation, which points to the changing *content* of the migration discourse. While the media coverage in the early stages of the refugee movements in 2015 focused on the tragic drownings in the Mediterranean sea, reasons for refugees' escape from their home countries, and humanitarian issues, the discourse slowly shifted to focus on perceived problems with the newly arrived. Especially new years eve 2015/16 marked a turning point, when hundreds of sexual abuses by men whom the victims perceived to be of arabic descent in Cologne were reported (if you look closely, you can see a slight discontinuity surrounding that date in the AfD-share, which increases around one and a half percentage points at that very date).

To be clear, I do not suggest that the salience of issues does not matter. Existing preferences in the electorate surely make certain issues more or less favourable for given parties at a specific moment in time. However, I suggest that the *content* of the public debate surrounding an issue shapes the mental representations of this issue in the electorate, and subsequently affects both their issue preferences and their voting behaviour. This would have profound implications for our understanding of party competition. For example, parties' strategic options look very different through this theoretical lens. Rather than remaining mute on unfortunate issues and accepting a fate determined by an uncontrollable media agenda, parties should communicate on any issue and promote arguments that bolster their position on this issue. Similarly, public preferences on an issue are not simply given *a priori*, but rather shaped by the considerations brought up in a changing public discourse.

How could changing issue content be described in a theoretical model? The next section introduces the concept of emphasis framing and the value-expectancy model of preference formation. This theoretical framework allows me to formulate clear expectations about the impact of changing issue content on voting behaviour, as well as clarify the concepts to be operationalised.

## The Value-Expectancy Framework and Emphasis Framing

What aspect of news content affect political preferences? Specifically, how can different considerations about the same issue affect one's reasoning and preferences regarding that issue? The large and established literature on framing has tackled precisely this question. To be clear, when I speak of "framing" in this manuscript, I refer to *emphasis framing*. Compared to concepts like equivalence frames (outlining different formulations of mathematically identical options; @Kahneman1979prospect, generic frames (portraying complex issues, e.g. elections within common and simplified metaphors, e.g. as a horse race; @DeVreese2005), or episodic frames (portraying issues through individual, personal stories; @Iyengar2005), emphasis framing focuses on the *emphasis* of different *considerations* surrounding an issue. For example, discussing the issue of migration, the debate can focus on the brutal reality of refugees' lifes and the hardships they face on their escape; or the discourse might focus on fears of the population in host countries and potential problems in the accommodation of refugees. The emphasis of different considerations will create different mental images which lead to differing assessments of the problem at hand, different preferences about the issue, and a different set of potential policy solutions.

More formally, this logic is expressed in the *value-expectancy framework* of attitude formation [@Ajzen2000; @Nelson1997Rally]. An attitude towards a political issue is a function of two things in this model: first, different considerations contain a certain positive or negative evaluation of the issue (let's call this $e_i$). For example, thinking of how to deal with criminal refugees will entail a more restrictive assessment towards immigration policy (simplistically $e_1 = -1$), whereas reasoning about how to prevent the humanitarian tragedies in the Mediterranean sea should be associated with a more welcoming position ($e_2 = 1$). All available considerations are then weighted by the *strength of the association* with the issue. That is, if an individual associates the issue of migration more strongly with criminal migrants rather than drownings in the Mediterranean sea, they weigh the former consideration more strongly ($w_1 = 0.9$) than the second ($w_2 = 0.1$) to form their overall opinion on the issue. The overall attitude $y$ is then a weighted sum of each consideration ($\sum_{i = 1}{e_i*w_i}$), or in this case $-1*0.9 + 1*0.1 = -0.8$. Given the stronger weighting of the consideration of criminal migrants by the respondent, the overall formed attitude is a negative/restrictive one.

## Emphasis Framing, the Media, and Party Competition

How does the media play into this? And how does this translate in to voting behaviour? The distinction of *frames in mind* and *frames in communication* is helpful [@Druckman2001]. *Frames in mind* describe the association mentioned above - a cognitive process where a given issue is associated with a specific consideration. *Frames in communication*, on the other hand, are located in political communication and are defined by the emphasis of one or several considerations in relation to an issue. Frames in communication change frames in mind by strengthening the association between an issue and a consideration, thereby affecting the cognitive definition of the issue and the resulting attitude towards this political issue. Hence, at least in theory, the media is able to change public opinion by emphasising different considerations about an issue.

How do changing issue attitudes translate into support for political parties? Similar to issue ownership theory, the theory here assumes that voters are more likely to incorporate a given issue in their voting decision if that issue is currently in public debate. However, dependent on the considerations that dominate public debate about this issue, parties' positions will be more or less favored. To stick with the migration example: while issue ownership theory assumes that the salience of migration will *always* favor radical right parties as the owner of that issue, the proposed theory suggests that which party profits from the increased salience of migration is also dependent on the considerations salient in the debate. If migrant crime is the dominating consideration, surely a radical right parties' restrictive positions will be supported and voters will be inclined to vote for this party. If however humanitarian issues dominate the debate, voters' cognitive representation of the issue should lead them to welcome more liberal positions on immigration policy, which should favor the radical right less and parties with more left-leaning positions more.

Lastly, I should mention that I assume that parties emphasise issue considerations in line with their political positions. That does not necessarily entail rationality (instead of a strategic effort to substantiate their positions, parties might also hold certain policy positions because they emphasize certain considerations), but is a (weak) assumption that should be mentioned here. Given that parties promote issue considerations supporting their positions, I can assume that when the news media emphasises considerations about a salient issue similar to that of the party, this party should experience increasing electoral support:

*Frame Competition Hypothesis: The more similar the media's framing of an issue to a given parties' issue framing, the better that parties' electoral prospect.*

One key difference compared to other models of party competition, especially the one described in @Meguid2005 is that in the present model, party positions are considered fixed, while it is the media debate that moves. Large shifts in parties' issue positions are generally rare [@SOURCE], but the media debate is in permanent flux, promoting a constantly changing set of issues and related considerations to voters. Political actors sponsor frames and arguments, only some of which get to influence the debate. While the classic model described in @Meguid2005 views parties' electoral performance as a function of parties position and issue salience, the present model views their electoral performance as a function of issue salience and the similarity of the dominant set of issue considerations to the set of considerations emphasized by the party to substantiate its position.

# Research Design

## The German *Bundestagswahl* 2017[^1]

[^1]: I am not entirely sure if it is smart to restrict myself to the 2017 campaign, as there should be more variation over time, especially moving from late 2015 to 2016, as described above. So far, this was a choice entirely made based on the use of the 2017 manifestos, but I might also combine the 2013 and 2017 manifestos and track the issue across time.

In order to study whether framing can change parties' electoral prospects, an ideal case would show substantial variation surrounding the framing of an issue but little variation in parties' positions. Ideally, the variation in salience would be limited and uncorrelated to the variation of the framing. I choose the case of the election campaign for the German Bundestagswahl 2017 as I consider most of these criteria are met. After the refugee movements throughout Europe in 2015, voters considered migration the most important issue [@Wessels2021, p. 12], and the media still discussed the issue, but at lower levels compared to 2017 (see figure \@ref(fig:salplot)). I expect that events during the election campaign, such as migrant crimes, conflicts regarding the intra-EU distribution of refugees, international terrorism, humanitarian crises as those among the Rohingya in Myanmar, as well as drownings in the Mediterranean sea lead to substantial variation in the considerations regarding migration being discussed - the key independent variable.

```{r salplot}

fread(here("data/media/bert_crime_daily.csv")) %>% 
  select(date_clean, mig_share) %>% 
  mutate(date_new = lubridate::floor_date(date_clean, "month")) %>% 
  group_by(date_new) %>% 
  summarise_all(mean) %>% 
  filter(date_new < as.Date('2018-01-01'),
         date_new >= as.Date("2014-01-01")) %>%
  ggplot(aes(x = date_new, y = mig_share)) +
  geom_line() + geom_point() +
  geom_vline(xintercept = as.Date("2017-09-24"), col = "red", lty = 2) +
  ggtitle('Migration salience in print media', '2014-2017, red line indicates 2017 election') +
  xlab('') + ylab('Share of migration articles') +
  theme_minimal()

```

## Measuring Migration Framing in the Media

How can such a diffuse concept as issue framing be measured? In order to identify different considerations in media coverage, I will use a two-step approach. After training a neural network (BERT) to classify migration content in news coverage (providing a measure of salience), I employ the structural topic model[^2] [@Roberts2019], to identify different topics being raised *within* migration coverage. This model allows to specify covariates, such as different prevalence of frames across newspapers. The topic model might for example identify topics like crime, terrorism, refugee deaths, and the distribution of refugees. Each of these topics represents a consideration about migration, which will have implications for the evaluation of the issue.

[^2]: At this point, I am not sure if I might not instead use a simple LDA.

The topic model then returns a topic vector for each news article, representing the distribution of topics in each article. For example, an article might talk to 90% about crime and 10% about the distribution of refugees, whereas it does not mention refugee deaths or terrorism. In this hypothetical four-topic space, the article would then be represented by a $[0.9, 0.1, 0, 0]$ vector, much like an embedding.

## Measuring Frame Similarity

In order to avoid making assumptions about the favorability of different considerations for different parties, I exploit this embedding-like structure of topic vectors to assess their similarity to party communication. The assumption I make instead is that party positions on migration are stable across the campaign, and that parties emphasize considerations supporting their position on the issue of migration. To measure the emphasis of different considerations in party communication, I select all paragraphs from parties' election manifestos in the 2017 election that are concerned with migration. For each party, I generate a topic vector using the model described in section, based on the collection of all relevant paragraphs \@ref(measuring-migration-framing-in-the-media).

Now, using these party-framing vectors, *I can assess the similarity of each news article towards each party's migration framing, simply by taking the cosine distance to the party's vector*. This way, it is possible to track the similarity of the media framing of migration to each party, in order to test the frame competition hypothesis: whether parties perform better when the media framing of migration is more similar to their own framing. This could well be done with document embeddings as well, however topic vectors offer a key advantage here: other than the dimensions of a, say, 300-dimensional embedding model, the dimensions of topic-vectors are interpretable. This can serve as a sanity check (e.g. if increased attention to crime results in higher similarity to a left-wing party that supports extended immigration, this would suggest that the model is not great), as well as enrich the data analysis: if frame similarity has an effect on parties' electoral prospects, it is easy to see which topics drive this effect and how the media debate affected opinion polls.

## Estimating the Effect of Frame Similarity on Electoral Prospects

As the above approach will provide me with time-series data on the similarity of the public debate surrounding migration to each parties' definition of migration, and given the time-series structure of the polling data (the dependent variable), it seems straightforward to use time-series modeling, and specifically vector-auto-regression (VAR) models[^3] [@Sims1980] to account for possible effects of the polls on media reporting about migration.

[^3]: As I never used VAR models to model interaction effects, I am not entirely sure how to account for variation in salience here. It might still be possible to use split samples of periods with higher and lower salience, or restrict the analysis to a period with relatively stable attention to migration (e.g. just the election campaign).

## Instrumenting Frame Attention

An additional possibility in order to improve causal identification was to use exogenous events like the new years eve in cologne, terrorist attacks abroad or drownings in the Mediterranean sea to instrument attention to specific topics. This idea has not far progressed, so I am curious about your thoughts on this.