### 2022-07-13 pres/Presentation.Rmd
Finish slides for OX|BER

---
title: "Frame Competition"
subtitle: "How News Framing affects Party Competition"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    seal: false
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
    includes:
      after_body: insert-logo.html
---

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#004A95", background_color = "#bfbfbf",
  base_font_size = "40px",
  header_h1_font_size = "1.3rem",
  header_h2_font_size = "1.2rem",
  header_h3_font_size = "1rem",
  footnote_font_size = "0.75rem",
  header_font_google = google_font("Lato", '700'),
  text_font_google   = google_font("Lato", "400"),
  code_font_google   = google_font("Fira Mono")
)

```

```{r, load_refs, include=FALSE, cache=FALSE}
library(RefManageR)
BibOptions(check.entries = FALSE, 
           bib.style = "authoryear",
           cite.style = "authoryear",
           style = "markdown",
           hyperlink = FALSE,
           max.names = 2, 
           longnamesfirst = F,
           dashed = FALSE)
myBib <- ReadBib("C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib", check = FALSE)
```

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = FALSE, 
  message = F, 
  warning = F, 
  cache = T, 
  out.width="90%", 
  fig.height=8, 
  fig.width=15, 
  fig.align='center', 
  eval = T, 
  dev = "svg"
  )
# knitr::opts_knit$set(root.dir = 'C:/Users/nicol/Dropbox/PhD/Papers/Bild/github/bild')

library(tidyverse)
library(dplyr)
library(data.table)
library(fixest)
library(lubridate)
library(patchwork)
library(here)

# load self-made functions
for (filename in list.files(here("code/functions"))){
  source(here(paste0('code/functions/', filename)))
}

```

class: inverse, center, middle

# Frame Competition

#### Nicolai Berk<sup>*</sup>

Presentation prepared for the first OX|BER Graduate Workshop

Berlin, 14th of July

.left[.footnote[<sup>*</sup> Dynamics RTG & Humboldt Universität Berlin]]

---

## Issue Ownership Theory

```{r IssueOwnership}

data.frame(
  Party = c("A", "A", "B", "B"),
  Issue = c("A's", "B's", "A's", "B's"),
  Votes = c(0.3, 0.1, 0.1, 0.3)
) %>% 
  ggplot(aes(Issue, Votes, col = Party)) +
  geom_point(size = 4) +
  geom_line(aes(group = Party), size = 2) +
  theme_xaringan() +
  xlab("Salient Issue")

```

---

## Parties behave accordingly

```{r MPAtt}

fread("https://www.dropbox.com/s/3taauo34rumthtm/final_dataset.csv?dl=1") %>% 
  mutate(date = as.Date(date, format = "%d/%m/%Y"),
         month = lubridate::floor_date(date, "quarter")) %>% 
  filter(date >= as.Date("2013-01-01")) %>% 
  group_by(month, sender) %>% 
  summarise(imm_att = mean(coding)) %>% 
  ggplot(aes(month, imm_att, fill = sender)) +
  geom_col(position = "dodge2") + 
  theme_minimal() +
  ggtitle("Attention to migration in mainstream parties' press releases", "2013-2019") +
  scale_fill_manual(values = c("#000000", "#e3000f")) +
  theme_xaringan() +
  ylab("") + xlab("")

```

---

## Research Question

- Parties only option to **remain mute** on other parties' issues?
- Framing literature suggests that issue definitions affect voters' policy preferences.
- Experimental work shows that framing can affect issue ownership (Seeberg 2019).

--

### Does framing affect voting behaviour?

---

class: inverse

## Argument

<br>

### 1. Issue framing affects issue attitudes.
### 2. Dominant media framing changes public opinion.
### 3. Policy preferences translate into voting behaviour.

--

## Media Framing $\rightarrow$ Voting Behaviour

---

## Research Design

<br>

### 1. Measure similarity of media framing to party framing.
### 2. Estimate effect of similarity on electoral performance.

---

## Data

### Media Content

Collected **2.5M news articles** from major German newspapers' websites (2013-2019).

.pull-left[

### (Expected) Electoral Performance

]

.pull-right[

  ![](https://www.politico.eu/wp-content/uploads/2022/02/01/Poll-of-Polls.svg)

]

- Polling data collected by *Politico's [Poll of Polls](https://www.politico.eu/europe-poll-of-polls/)*.
- Generates daily moving average of party support.

---

## Data II

<br>

### Party framing

- 2017 Manifesto corpora, reduced to migration content.
- Ideal because **summarise general position** on the issue.
- Downside: assumption of stable position over time.

---

## Measuring Issue Framing

.pull-left[

  ![](vis/CognitiveStorage.png)

]

.pull-right[

1. Identify media content relating to issue:
  - Annotate sample
  - Supervised ML (BERT)
2. Identify considerations:
  -  Topic model

]

---

## Measuring Frame Similarity

```{r vectors, fig.height=7, cache=FALSE}

data.frame(
  item = c("News Article", "Party A", "Party B",
            "News Article", "Party A", "Party B"),
  dim1 = c(0,  0.2, 0.8,
            0,0,0),
  dim2 = c(1, 0.8, 0.2,
            0,0,0)
) %>% 
ggplot(aes(x = dim1, y = dim2, 
        color = item, linetype = item)) +
  geom_line(size = 2) +
  geom_point(alpha = 0.3, size = 4) + 
  theme_minimal() +
  xlab("Attention to Consideration 1") +
  ylab("Attention to Consideration 2") +
  theme(legend.title = element_blank()) +
  theme_xaringan()

```

---

## Measuring Frame Similarity

### Why frame *similarity*?

- Parties communicate a certain framing either as cause or outcome of their positions.
- If media framing is **similar**, this supports a parties' position.
- As a result, citizens are more likely to vote for party.

--

### Measurement

$$Similarity_{Party, t} = cos(TopicVector_{Media}, TopicVector_{Party})_t$$

---

## SVAR Time Series Model

<br>

- VAR allows modeling multiple time series as functions of own and other variables past.
- Structural VAR allows modeling interaction $Performance = Salience*Similarity$.
- Estimate of interest: effect of similarity on electoral performance.

---

## Instrumental Variable Approach

.left-column[

```{r dagInstrument, fig.align='center'}
DiagrammeR::grViz("
digraph boxes_and_circles {

  rankdir = 'TB'
  bgcolor = '#bfbfbf'

  # a 'graph' statement
  graph [overlap = true, 
          fontsize = 16]

  # several 'node' statements
  node [shape = plaintext,
        fontname = Verdana,
        fontcolor = black]
  Z [label = 'Events', fontcolor = '#004A95']
  X [label = 'Media Framing']
  Y [label = 'Voting']

  # several 'edge' statements
  Z->X [color = '#004A95']
  X->Y [color = black]
}
")

```

]

.right-column[

Events drive media attention to considerations, e.g.

- Migrant crime,
- Terrorist Attacks,
- Drownings.

### These should only affect public opinion through media coverage.

]

---

## Instrumental Variable Approach II

<br>

- Framing can be instrumented using exogenous events.
- Predicted frame attention can then be used to estimate causal effect (**2SLS**).

---

## Summary

<br>

- **Goal**: Project tries to challenge assumption that issues always affect voters similarly.
- **Argument**: dominant media framing of issues affects party support.
- **Methods**: 
  + innovative measure of frame similarity,
  + combination of time-series and causal inference tools (SVAR + IV).

---

## Open Questions

- Does theory make sense?
- Time Frame (2017 campaign or longer?)
- Manifestos for party frames?
- SVAR correct model? (might go for unidirectional model with 2SLS)
- Additional Experiment?
- ...?

---

class: inverse, middle, center

# Fin

## Thank you!

.pull-left[![](vis/banner_new.png)]

.pull-right[![](vis/dfg_logo_englisch_blau_en.jpg)]