### 2023-05-04 pres/Presentation.qmd
Presentation Dynamics

---
title: "Does Media Framing Affect Party Performance?"
subtitle: "Presentation at the Dynamics RTG Colloquium"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
date-format: "DD.MM.YYYY"
logo: vis/dynamics.jpg
footer: "Nicolai Berk | Media Framing and Party Performance"
format: 
  revealjs:
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
bibliography: "C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib"
---

## Motivation
```{r setup, echo=F, include=F}

library(tidyverse)
library(dplyr)

```

![Source: [Pew Research Center](https://www.pewresearch.org/fact-tank/2022/09/08/republicans-and-democrats-have-different-top-priorities-for-u-s-immigration-policy/)](vis/pewRD.png)

## Emphasis Framing

<br>

- Voters draw on different <mark>considerations</mark> when evaluating issues.
- Exposure to different considerations changes attitudes. <font size="5"> [@Nelson1997Rally]</font>
- Spatial model suggests this should translate into voting behaviour.
- Party competition research mostly concerned with issue emphasis. <font size="5"> [@GreenPedersen2019]</font>

## Contradicting Predictions

<br>

::: columns

::: {.column}

### Issue Ownership

```{mermaid}
%%{
  init: {
    'theme': 'base',
    'themeVariables': {
      'primaryColor': '#003366',
      'primaryTextColor': '#fff',
      'primaryBorderColor': '#003366',
      'lineColor': '#2171c2',
      'secondaryColor': '#2171c2',
      'tertiaryColor': '#fff'
    }
  }
}%%
flowchart LR
  A(Issue\nSalience) --> B(Party\nPosition)
  B --> C(Voting)
```

*All forms of immigration coverage benefit RR*

:::
::: {.column}

### Emphasis Framing
```{mermaid}
%%{
  init: {
    'theme': 'base',
    'themeVariables': {
      'primaryColor': '#003366',
      'primaryTextColor': '#fff',
      'primaryBorderColor': '#003366',
      'lineColor': '#2171c2',
      'secondaryColor': '#2171c2',
      'tertiaryColor': '#fff'
    }
  }
}%%
flowchart LR
  A(Frame\nPrevalence) --> B(Voter\nPosition)
  B --> C(Voting)
```

*Pro-Immigration Frames hurt RR*

:::
:::

# Empirical Strategy

## Measuring Media Framing {background-image="https://raw.githubusercontent.com/ddangelov/Top2Vec/master/images/hdbscan_docs.png" background-size="70%" background-repeat="no-repeat" background-opacity="0.3" background-position="right 0% bottom 0%"}

<br>

1. Identify migration content with supervised [BERT model](#BERT).
2. Generate <mark>document embeddings</mark> for migration articles.
3. Cluster embeddings into topics using `top2vec`.
4. Annotate based on most similar document and word vectors.

## Validation of Frame Estimates

![Different Crime Measures](../plots/Measures/crime_estimates.png)

## Validation Frame Estimates II

:::: {.columns}
::: {.column width="50%"}

![Document Level](../plots/Measures/crime_est_cor_doclevel.png)

:::

::: {.column width="50%"}

![Monthly Aggregation](../plots/Measures/crime_est_cor_monthly.png)

:::
::::

## Topic Distribution over Time

![](../plots/descriptives/topics_descriptives.png)

## Measuring Party Preferences

<br><br>

- Current: Aggregated polling from [Politico's PollOfPolls](https://www.politico.eu/europe-poll-of-polls/).
- Planned: Daily [forsa-Bus](https://search.gesis.org/?source=%7B%22query%22%3A%7B%22bool%22%3A%7B%22must%22%3A%7B%22match_all%22%3A%7B%7D%7D%2C%22filter%22%3A%5B%7B%22term%22%3A%7B%22type%22%3A%22research_data%22%7D%7D%2C%7B%22term%22%3A%7B%22study_group.keyword%22%3A%22Forsa-Bus%22%7D%7D%5D%7D%7D%7D&lang=en) polls of sample of 500 German voters.

## Party Performance

![](../plots/descriptives/polls.png){fig-align="center"}

## Estimating Effects on Party Performance

<br>

#### Inferential: RDD

Exploit events creating <mark>discontinuities</mark> in migration coverage.

<br>

#### Descriptive: VAR

Assess <mark>over-time correlation</mark> of frames and party polls.

# Preliminary Results

# Regression Discontinuity Design

## RDD

## VAR

![](../plots/var/top_topics_weekly.png)

# Conclusion

## Preliminary Summary of Results

<br><br>

- Framing *does* affect polling.
- But unclear how this generalises.

## Next Steps

<br><br>

- Get better data on dependent variable.
  - Clear RDD estimate.
  - Fine-grained VAR.
- Look for additional events.
- More validation of topics.

# Thank you!

<br>

This research is generously supported by:

::: {layout-ncol=2}

![](vis/banner_new.png)

![](vis/dfg_logo_englisch_blau_en.jpg)

:::

# Resources {visibility="uncounted"}

::: {#refs}
:::

# Appendix {visibility="uncounted"}

## Dictionary Estimates

![](../plots/var/top_dicts_weekly.png)

## Why [BERT](https://huggingface.co/docs/transformers/model_doc/bert) {#bert visibility="uncounted"}

<br>

- Pre-trained on large language corpus. <font size="5">([Details](https://www.deepset.ai/german-bert))</font> \
<font color='#801414'>➡ less training data, better performance.</font>
- Uses **contextual** embeddings. \
<font color='#801414'>➡ "fish & **chips**" vs "poker **chips**".</font>
- Outperforms humans on [SQuAD](https://rajpurkar.github.io/SQuAD-explorer/).
- SOTA in text classification.

<br>

[BACK](#measuring-media-framing)

![](vis/bertarch.png){.absolute width=400 bottom=0 right=0}

## Dealing with Imbalance {#imbalance visibility="uncounted"}

::: columns
::: {.column width="70%"}

::: callout-important
## Imbalance: Little Information in Random Samples
:::

::: callout-tip
## Informed Approach

- Generate dictionary related to outcome class.
- Extend using word embeddings to minimize bias.
- Count occurences in data.
- Draw stratified sample for subsequent training.

**Ensures overweighting of minority class.**

:::
:::

::: {.column width="30%"}

<br>

```{r, fig.height=4, fig.width=3}

counts <- 
  round(rbeta(1000, 0.2, 0.5)* 10, 0) %>%
  as_tibble() %>% 
  mutate(sample_id = cut(value, 
                         c(-Inf, 0.5, 2.5, Inf), 
                         c(1, 2, 3)),
         id = 1:n())

counts$sample_T <- rbinom(nrow(counts), 1, 0.05*(counts$value+1))

counts %>% 
  ggplot(aes(x = value)) +
  geom_histogram(aes(fill = as.character(sample_T)), bins=3) +
  scale_fill_manual(values = c("gray", "#003366"),
                    labels = c("not sampled", "sampled")) +
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  xlab("# of dict terms in text") + ylab("Observations")

```

:::
:::

[BACK](#mediadata)