### 2023-06-14 pres/Presentation.qmd
move to quarto

---
title: "Framing and Voting"
subtitle: "CompText 2023"
author: "Nicolai Berk"
institute: "Humboldt Universität Berlin"
date: "`r Sys.Date()`"
date-format: "DD.MM.YYYY"
logo: vis/dynamics.jpg
footer: "Nicolai Berk | Framing and Voting"
format: 
  revealjs:
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
bibliography: "C:/Users/nicol/Dropbox/Studium/BibTex/PhD.bib"
---

# How Does Media Content Affect Voting Behaviour?

## Issue Ownership
```{r setup, echo=F, include=F}

library(tidyverse)
library(dplyr)

```

::: {.columns}
::: {.column width="50%"}

![Source: [Conservatives Website](https://www.conservatives.com/)](vis/conservatives.png)

:::
::: {.column width="50%"}

![Source: [Labour Website](https://shop.labour.org.uk/category/leaflets)](vis/LabourNHS.png)

:::
:::

## Emphasis Framing

```{r frames, out.width="60%"}

library(patchwork)

dta <- 
  data.frame(
    id = 1:100,
    pro_frame = rnorm(10000,  1),
    con_frame = rnorm(10000, -1)
  )

p1 <- 
  dta %>%
  ggplot(aes(con_frame)) +
  geom_histogram(fill = "#e3003a") +
  theme_minimal() +
  labs(
    title = "Left-leaning Frame",
    x = "", y = ""
  ) +
  xlim(c(-2, 2))

p2 <- 
  dta %>%
  ggplot(aes(pro_frame)) +
  geom_histogram(fill = "#00b7e9") +
  theme_minimal() +
  labs(
    title = "Right-leaning Frame",
    x = "Issue Position", y = ""
  ) +
  xlim(c(-2, 2))

p1/p2

```
<!-- 
## Contradicting Predictions

<br>

### Issue Ownership

```{mermaid}
%%| fig-width: 10
%%{
  init: {
    'theme': 'base',
    'themeVariables': {
      'primaryColor': '#003366',
      'primaryTextColor': '#fff',
      'primaryBorderColor': '#003366',
      'lineColor': '#2171c2',
      'secondaryColor': '#2171c2',
      'tertiaryColor': '#fff'
    }
  }
}%%
flowchart LR
  A(Issue\nSalience) --> B(Party\nPosition)
  B --> C(Voting)
```

*All forms of immigration coverage benefit RR*

## Contradicting Predictions

<br>

### Emphasis Framing
```{mermaid}
%%| fig-width: 10
%%{
  init: {
    'theme': 'base',
    'themeVariables': {
      'primaryColor': '#003366',
      'primaryTextColor': '#fff',
      'primaryBorderColor': '#003366',
      'lineColor': '#2171c2',
      'secondaryColor': '#2171c2',
      'tertiaryColor': '#fff'
    }
  }
}%%
flowchart LR
  A(Frame\nPrevalence) --> B(Voter\nPosition)
  B --> C(Voting)
```

*Pro-Immigration Frames hurt RR*
 -->

# Empirical Strategy

## Measuring Media Framing {background-image="https://raw.githubusercontent.com/ddangelov/Top2Vec/master/images/hdbscan_docs.png" background-size="70%" background-repeat="no-repeat" background-opacity="0.3" background-position="right 0% bottom 0%"}

<br>

1. Collect 2.5M News Articles.
2. Identify migration content with supervised [BERT model](#BERT).
3. Generate <mark>document embeddings</mark> for migration articles.
4. Cluster embeddings into topics using `top2vec`.

## BERT Performance

<br>

| Metric | Score |
| :------------- |  ---- |
| **Accuracy**       | 0.96  |
| **Precision**      | 0.94  |
| **Recall**         | 0.93  |
| **F1-Score**       | 0.94  |

<!-- ## Validation of Frame Estimates

![Different Crime Measures](../plots/Measures/crime_estimates.png) -->

## Validation Frame Estimates

:::: {.columns}
::: {.column width="50%"}

![Document Level](../plots/Measures/crime_est_cor_doclevel.png)

:::

::: {.column width="50%"}

![Monthly Aggregation](../plots/Measures/crime_est_cor_monthly.png)

:::
::::

## Topic Distribution over Time

![](../plots/descriptives/topics_descriptives.png)
<!-- 
## Measuring Party Preferences

<br><br>

- Current: Aggregated polling from [Politico's PollOfPolls](https://www.politico.eu/europe-poll-of-polls/).
- Planned: Daily [forsa-Bus](https://search.gesis.org/?source=%7B%22query%22%3A%7B%22bool%22%3A%7B%22must%22%3A%7B%22match_all%22%3A%7B%7D%7D%2C%22filter%22%3A%5B%7B%22term%22%3A%7B%22type%22%3A%22research_data%22%7D%7D%2C%7B%22term%22%3A%7B%22study_group.keyword%22%3A%22Forsa-Bus%22%7D%7D%5D%7D%7D%7D&lang=en) polls of sample of 500 German voters. -->

## Party Performance

![](../plots/descriptives/polls.png){fig-align="center"}

## Estimating Effects on Party Performance

<br>

#### Inferential: RDD

Exploit events creating <mark>discontinuities</mark> in migration coverage.

<br>

#### Descriptive: VAR

Assess <mark>over-time correlation</mark> of frames and party polls.

# Preliminary Results

## Tragedy off Lampedusa: Media Effect

![](../plots/RDD/MediaEffect_lampedusa_med.png)

## Tragedy off Lampedusa: Media Effect

![](../plots/RDD/PollEffect_lampedusa.png)

## Parndorf Incident: Media Effect

![](../plots/RDD/MediaEffect_parndorf_ht.png)

## Parndorf Incident: Polling Effect

![](../plots/RDD/PollEffect_parndorf.png)

## VAR

![](../plots/var/top_topics_weekly.png)

## Next Steps

<br><br>

- Currently waiting for daily opinion polls from [forsa-Bus](https://search.gesis.org/?source=%7B%22query%22%3A%7B%22bool%22%3A%7B%22must%22%3A%7B%22match_all%22%3A%7B%7D%7D%2C%22filter%22%3A%5B%7B%22term%22%3A%7B%22type%22%3A%22research_data%22%7D%7D%2C%7B%22term%22%3A%7B%22study_group.keyword%22%3A%22Forsa-Bus%22%7D%7D%5D%7D%7D%7D&lang=en).
- Add control variables (esp. immigration numbers).

# Thank you!

<br>

This research is generously supported by:

::: {layout-ncol=2}

![](vis/banner_new.png)

![](vis/dfg_logo_englisch_blau_en.jpg)

:::